Pop-up cycleways: a peak into post COVID-19 transport futures

### 2020-05-11 conversation.Rmd
Update conversation article - work in progress

Around the world, as nations prepare to emerge from months of COVID-19 induced lockdown, life cannot simply return to previous norms. The pandemic has transformed all sectors of the economy, not least transport. Long distance travel has been dramatically altered, as [demand plummets](https://osf.io/preprints/socarxiv/v3g5d/) and [airline companies go bust](https://arxiv.org/abs/2004.08460). Yet short distance travel patterns have also changed greatly. Forced to stay close to home, many people have started walking and cycling more. From a health and well-being perspective, this may be a more important shift, especially in the long term as lockdown measures are gradually eased.

Active travel is already [recommended by the World Health Organisation](https://who.canto.global/pdfviewer/viewer/viewer.html?share=share%2Calbum%2CU6GDM&column=document&id=q3v02qdsh114763ffe3apo4s2p&suffix=pdf) as a safe mode of transport, and in the coming weeks and months it will play a vital role in enabling the reopening of many countries. As travel increases, we must ensure that safe, socially distanced walking and cycling remains possible. Maintaining social distancing on public transport could severely reduce the capacity of rail and bus networks, and there is a risk that this may lead to a rapid rise in road traffic.  [ record bicycle sales ] suggest that many people have taken advantage of lockdown conditions to take up or return to cycling, but without careful management, increased traffic congestion could again deter many would-be cyclists from the streets.

Active travel has extensive [ social and environmental benefits ]. Physical activity is associated with XX health conditions….. mental health…..isolation...

Evidence is beginning to emerge of [ potential links ] between air pollution and COVID-19 lethality, not least because many health conditions such as asthma and xxx are [ worsened by poor air quality ]. Indeed, it is estimated that global improvements in air quality in recent months have led to [ xx thousand fewer deaths ] from conditions such as these. This makes it all the more important to restrict any rise in road traffic and to encourage more beneficial means of travel.

Local and national governments worldwide are still working out how best to respond to these challenges. On Saturday 9th May, British Transport Secretary Grant Shapps announced a £2 billion investment in measures to promote walking and cycling. This follows a £10 million commitment from the Scottish Government and action announced in Wales. Hundreds of kilometres of temporary infrastructure is being put in place in dozens of cities worldwide, from [Milan] to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf). Never has there been such a need for affordable, healthy and socially distanced transport that does not create the fumes that [emerging evidence] links to coronavirus susceptibility.

In terms of practical steps to encourage walking and cycling, a range of simple interventions can be undertaken, initially on a temporary basis. Roads can be closed to motor vehicles, or physical barriers can be used to enable [ filtered permeability ], in which access for residents is retained but modal filters enable routes for walking and cycling that are faster and more direct than the routes for motor vehicles. This is well established in [ the London Borough of Hackney ] where attractive barriers such as planters have been used to good effect. Another option involves keeping the roads open while reallocating excess roadspace to widen pavements or introduce pop-up cycleways.

@@

Building on the previous projects such as the [ Propensity to Cycle Tool ] and the [ Cycling Infrastructure Prioritisation Toolkit ], researchers from the University of Leeds have investigated where roadspace could be reallocated for pop-up cycleways. They have mapped out the top locations for intervention in ten major cities across England, with the intention of rolling out these methods across the country in the near future.

>>>>

Around the world, as nations prepare to emerge from months of COVID-19 induced lockdown, life cannot simply return to pre-pandemic normal. The pandemic has transformed all sectors of the economy, not least transport, with much attention on [plummeting levels of long distance travel](https://osf.io/preprints/socarxiv/v3g5d/) and [airline companies going bust](https://arxiv.org/abs/2004.08460). Yet short distance travel patterns have also changed greatly and, perhaps, forever.
This may be a more important shift from a health perspective, especially in the long term as lockdown measures are gradually eased.

Active travel is already [recommended by the World Health Organisation](https://who.canto.global/pdfviewer/viewer/viewer.html?share=share%2Calbum%2CU6GDM&column=document&id=q3v02qdsh114763ffe3apo4s2p&suffix=pdf) during the outbreak.
With health and scientific evidence at the top of government priorities, walking and cycling will play a vital role in enabling the reopening of many countries.
There are two main reasons why walking and cycling levels can be expected to grow during the pandemic and its aftermath. First, forced to stay close to home, many people have started walking and cycling more for health reasons. Second, public transport options have been greatly reduced in many [countries](https://doi.org/10.1126/science.abb6105), voluntary use has fallen, and services that do run will run at a fraction of their intentended capacity. 
As UK Transport Minister Grant Shapps [said](https://www.gov.uk/government/speeches/transport-secretarys-statement-on-coronavirus-covid-19-9-may-2020) on 9th March, "once you take into account the 2 metre social distancing rule, there would only be effective capacity for one in ten passengers on many parts of the network."
This second reason is particularly important given that many 'key workers' who keep economies and vital services moving are low paid, reliant on public transport, and with limited capacity to buy and run private automobiles.

Furthermore, evidence is beginning to emerge of [ potential links ] between air pollution and COVID-19 lethality, not least because many health conditions such as asthma and xxx are [ worsened by poor air quality ].
Indeed, it is estimated that global improvements in air quality in recent months have led to [ xx thousand fewer deaths ] from conditions such as these.
This makes it all the more important to restrict any rise in road traffic and to encourage more beneficial means of travel.

To some extend the authorities are responding to bottom-up demand, with [sales of bicycles have gone through the roof](https://www.bbc.co.uk/news/business-52564351) suggesting that there is a huge health opportunity.
There is a [window of opportunity](link to Forbes article) that could be missed if roads become busy and dangerous again before new active transport habits become embedded.

<!-- Active travel has extensive [ social and environmental benefits ]. Physical activity is associated with XX health conditions….. mental health…..isolation... -->
In response to this emerging evidence, local and national transport authorities have been scrambling to ensure that walking and cycling levels can grow, while social distancing guidance is observed.
A range of interventions has been [proposed](link to academic article from USA - rl).
On Saturday 9th May, British Transport Secretary Grant Shapps announced a £2 billion investment in measures to promote walking and cycling. This follows a £10 million commitment from the Scottish Government and action announced in Wales. Hundreds of kilometres of temporary infrastructure is being put in place in dozens of cities worldwide, from [Milan] to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf). Never has there been such a need for affordable, healthy and socially distanced transport that does not create the fumes that [emerging evidence] links to coronavirus susceptibility.

In terms of practical steps to encourage walking and cycling, a range of simple interventions can be undertaken, initially on a temporary basis. Roads can be closed to motor vehicles, or physical barriers can be used to enable [ filtered permeability ], in which access for residents is retained but modal filters enable routes for walking and cycling that are faster and more direct than the routes for motor vehicles. This is well established in [ the London Borough of Hackney ] where attractive barriers such as planters have been used to good effect. Another option involves keeping the roads open while reallocating excess road space to widen pavements or introduce pop-up cycleways.

@@

Building on the previous projects such as the [ Propensity to Cycle Tool ] and the [ Cycling Infrastructure Prioritisation Toolkit ], researchers from the University of Leeds have investigated where road space could be reallocated for pop-up cycleways. They have mapped out the top locations for intervention in ten major cities across England, with the intention of rolling out these methods across the country in the near future.
